---
title: "zzysdata Package"
output: github_document
---

## zzysdata Package

The **zzysdata** package provides a simple example of creating a custom R package with data visualization and a Shiny application.\
It includes: - A dataset (`clean_data`) - A custom plotting function (`plot_scatter()`) - A Shiny app for interactive visualization

### Installation and load

You can install the package (if not already installed) using:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Install zzysdata if it’s not yet installed
if (!requireNamespace("zzysdata", quietly = TRUE)) {
  devtools::install(build = TRUE, reload = TRUE)
}

library(zzysdata)
library(ggplot2)

```

### Overview

The zzysdata package contains example datasets and functions for data visualization and analysis. It was developed as part of an R package assignment to demonstrate dataset creation, function documentation, and plotting.

### Data

The package includes a dataset called clean_data which contains site and variable information for environmental analysis.

```{r}
data("clean_data")
head(clean_data)
```

### Visualization

We can also visualize the clean_data dataset using ggplot2.

```{r}
ggplot(clean_data, aes(x = Site, y = Variance, fill = Variable)) +
geom_bar(stat = "identity", position = position_dodge()) +
labs(
title = "Relative Importance of Predictors Across Sites",
x = "Stream Site",
y = "Variance Explained (%)"
) +
theme_minimal()

```

### Project Summary
This project involved developing an R package named zzysdata as part of the ETC5523 “Communicating with Data” Assignment 4. The package contains a cleaned dataset, useful data exploration functions, and an interactive Shiny application. Using devtools and usethis, I structured the package with directories for raw data, documentation, and the Shiny app. The function plot_scatter() was created to visualize relationships between variables and tested successfully using the mtcars dataset. The Shiny app, located in inst/app/, allows users to interactively explore the data and visualize trends dynamically. All components, including documentation generated with roxygen2, were verified through devtools::document(), build(), and check(). Finally, a pkgdown site was configured to host the package documentation and vignettes. Overall, this project demonstrates the complete workflow from data cleaning and analysis to creating a user-friendly, interactive, and well-documented R package.
